---
title: "mapvizier Class Construction"
author: "Chris Haid"
date: "June 4, 2014"
output: html_document
---

## Motivation
`mapvizier` is classy and so I think it deserves a class object that we can write our functions against. We will use the *S3* object oriented method dispatch in `R`, which you can learn about [from Hadley Whickham](http://adv-r.had.co.nz/OO-essentials.html)


```{r packaes_and_data}
require(mapvisuals)
require(dplyr) # make sure you have the latest version
data(nweamap)
glimpse(nweamap)
```

I first want to define a function that separates out term names as per our earlier email and calculates cohort names (i.e., the SY 2013-14 8th grades are in 2018), since it makes sense to do this for the mapvizier class objects:

# Term Split Function
```{r term_split}
term_split2 <- function(term_name, grade=NULL){
  
  require(stringr)
  # extract Season
  Season<-str_extract(term_name, 
                      "[[:alpha:]]+"
                      )
  # extract Year1
  Year1<-as.integer(str_extract(term_name, 
                                "[[:digit:]]+"
                                )
                    )
  
  # extract Year2
  Year2<-as.integer(gsub("([a-zA-Z]+[[:space:]][[:digit:]]+-)([[:digit:]]+)",
                         "\\2", 
                         term_name
                         )
                    )
  # construct School Year (e.g, "2012-2013")
  SY<-paste(Year1, Year2, sep="-")
  
  # calculate Cohort Year
  if(!is.null(grade)) {
    CohortYear<-(12-grade)+as.numeric(Year2)
    #return df
    x<-data.frame(Season, Year1, Year2, SY, CohortYear, stringsAsFactors = FALSE)
    x
   } else {
    x<-data.frame(Season, Year1, Year2, SY, stringsAsFactors = FALSE)
    x
  }
}
```

Here's a quick test of that function:
```{r test_term_split}
test_termsplit<-term_split2(term_name = nweamap$TermName)
glimpse(test_termsplit)
test_termsplit_with_grade<-term_split2(term_name = nweamap$TermName,
                           grade = nweamap$Grade
                           )
glimpse(test_termsplit_with_grade)

```


Cool, so that works!  Let's start by making a constructor function, which will in this first instance, simply add the `term_split2` columns to the `nweamap` data. Then I'll add some school abbreviations, KIPP's weird quartile calculation, and KIPP Tiered Growth multipliers (since these functions are already in mapvizier!).

### Creating a class
We will define a method that calls the constuctor function, does the calculations and appending and returns a new data.frame objectof the class `mapvizier`.

```{r class_constructor1}
# define a generic 
mapvizier <- function(x) UseMethod("mapvizier")

# now define the default method (you have to use the . here, which 
# is proably why you don't want to use dots (.) elsewhere even if 
# they are super duper easy to type, since no shifting is required)
mapvizier.default <- function(x){
  require(dplyr)
  require(data.table)
  # coerce to data.frame to be safe
  x<-as.data.frame(x)
  
  x2 <- cbind(x,
              term_split2(term_name = x$TermName,
                          grade = x$Grade)
              )
  
  # lookout!  here comes some dplyr sugar to get the munging done
  # Adding school abbreviations, TestQuartile, and KIPP Tiered Growth
  x3 <- x2 %>% filter(GrowthMeasureYN=="TRUE") %>%
     mutate(SchoolInitials   = abbrev(SchoolName), 
            TestQuartile     = kipp_quartile(TestPercentile),
            KIPPTieredGrowth = tiered_growth(TestQuartile, Grade)
         )
  
  # let's add some proper growth norms from the growth norms table
  
  map_data <- cbind(x3,
              nwea_growth(x3$Grade, 
                          x3$TestRITScore, 
                          x3$MeasurementScale
                          )
              )
  #shit, why not just to all the joins we could ever want on this original data
  # Create Seaason to Season Numbers
 year_list<-unique(map_data$Year2)

  map.SS<-rbindlist(lapply(year_list, 
                         s2s_match, 
                         .data=map_data, 
                         season1="Spring", 
                         season2="Spring", 
                         typical.growth=T,
                         college.ready=T
                         )
                  )
 map.FS<-rbindlist(lapply(year_list, 
                         s2s_match,
                         .data=map_data, 
                         season1="Fall", 
                         season2="Spring", 
                         typical.growth=T,
                          college.ready=T
                                 )
                          )
map.FW<-rbindlist(lapply(year_list, 
                         s2s_match, 
                        .data=map_data, 
                         season1="Fall", 
                         season2="Winter", 
                         typical.growth=T,
                         college.ready=T
                               )
                        )
map.WS<-rbindlist(lapply(year_list,
                         s2s_match, 
                         .data=map_data, 
                         season1="Winter", 
                         season2="Spring", 
                         typical.growth=T,
                         college.ready=T
                         )
                  )
map.FF<-rbindlist(lapply(year_list, 
                         s2s_match, 
                         .data=map_data, 
                         season1="Fall", 
                         season2="Fall", 
                         typical.growth=T,
                         college.ready=T
                         )
                  )

map.all.growth<-rbindlist(list(map.SS, map.FS, map.FW, map.WS, map.FF))
  
# add to mapviz object

mapviz<-list(mapData = map_data,
             seasonMatched = map.all.growth
             )

class(mapviz) <- "mapvizier"
  
mapviz            
}
```

ok, let's see if this fucker works:

```{r mapviz_class_test}
test_mv<-mapvizier(nweamap)
str(test_mv)
```

### Methods 
Sweet!!!! That works.  So now we can pass something that looks like `nweamap` and add some columns that we use in a lot of visualizations already and get the step of matching seasons out of the way.  So let's write a couple of methods against this class: `print()`, `summary()`, and `plot()` (NB: we will name these function `print.mapvizier`, `summary.mapvizier()`, and `plot.mapvizier()`, which allows `R` to use method dispatch when we call `print(x)` and `x` is `mapvizier` object).

#### print()
```{r print.mapvizier}
print.mapvizier <-  function(x, ...) {
  require(dplyr)
  
  #gather some summary stats
  n_sy <- length(unique(x$mapData$SY))
  min_sy <- min(x$mapData$SY)
  max_sy <- max(x$mapData$SY)
  n_students <- length(unique(x$mapData$StudentID))
  n_schools <- length(unique(x$mapData$SchoolName))
  growthseasons <- unique(x$seasonMatched$GrowthSeason)
  n_growthseasons <- length(growthseasons)
  
  cat("A mapvizier object repesenting:\n- ")
  cat(paste(n_sy))
  cat(" school years from SY")
  cat(paste(min_sy))
  cat(" to SY")
  cat(paste(max_sy))
  cat(";\n- ")
  cat(paste(n_students))
  cat(" students from ")
  cat(paste(n_schools))
  cat(" schools;\n- and, ")
  cat(paste(n_growthseasons))
  cat(" growth seasons:\n    ")
  cat(paste(growthseasons, collapse = ",\n    "))

  
  }

#I'm adding a glimpse method too, since it is cool and I want to see both attached dataframes is the mapvizier object

# create genereic
glimpse<-function(x) UseMethod("glimpse")

# assigne dplry's glimpse to glimpse.default to preserve glimpse defualt 
# behavrio
glimpse.default <- function(tbl, width) dplyr::glimpse(tbl, width=getOption("width"))

# Now create method for mapvizier class
glimpse.mapvizier <- function(tbl, width=getOption("width")){
  #require(dplry)
  cat("mapData:\n")
  print(dplyr::glimpse(tbl$mapData, width))
    cat("seasonMatched:\n")
  print(dplyr::glimpse(tbl$seasonMatched, width))
}
```
And now let's see if these work:

```{r test_print}
test_mv # this automatically calls the print method
print(test_mv)  # same as above

glimpse(test_mv)
```

Excelent! This is going great. Lets do a dead simple summary method:

#### summary()
```{r summary_method}

summary.mapvizier <- function(object, ..., digits){
  
  mapData<-group_by(as.data.frame(object$seasonMatched), 
                    SY.2, 
                    GrowthSeason, 
                    SchoolInitials, 
                    Grade.2, 
                    MeasurementScale
                    )
  mapSummary <- dplyr::summarize(mapData,
                          N = n(),
                          N_Typical = sum(MetTypical),
                          Pct_Typical = round(sum(MetTypical)/N,2),
                          N_CollegeReady = sum(MetCollegeReady),
                          Pct_CollegeReady = round(sum(MetCollegeReady)/N,2),
                          N_50th_Pctl_S1 = sum(TestPercentile>=50),
                          Pct_50th_Pctl_S1= round(sum(TestPercentile>=50)/N,2),
                          N_50th_Pctl_S2 = sum(TestPercentile.2>=50),
                          Pct_50th_Pctl_S2 = round(sum(TestPercentile.2>=50)/N,2),
                          N_75th_Pctl_S1 = sum(TestPercentile>=75),
                          Pct_75th_Pctl_S1 = round(sum(TestPercentile>=75)/N,2),
                          N_75th_Pctl_S2 = sum(TestPercentile.2>=75),
                        Pct_75th_Pctl_S2 = round(sum(TestPercentile.2>=75)/N,2)
  )

  
  setnames(mapSummary, 
         c("SchoolInitials", "Grade.2", "MeasurementScale", "SY.2"),
         c("School", "Grade", "Subject", "SY")
         )

#class(mapSummary)<-"mapvizierSummary"  
  
class(mapSummary)<-c("mapvizierSummary", class(mapSummary))

#return
mapSummary
  
}

```
And again, let's test this:

```{r test_summary_method}
test_mv_summary<-summary(test_mv)

test_mv_summary
```


#### plot()

Finally let's do a plot method on the summary object

```{r plot_method}
plot.mapvizierSummary <-  function(x, 
                                   growthseason="Fall - Spring", 
                                   subjects=c("Reading", 
                                              "Mathematics", 
                                              "General Science"),
                                   grades=c(2:8)){
  require(ggplot2)
  x<-as.data.frame(x)
  plot_data <- filter(x, 
                      GrowthSeason == growthseason,
                      Subject %in% subjects,
                      Grade %in% grades
                      )
  
  p.long<-ggplot(plot_data, 
                 aes(x=gsub("20","",SY), 
                     y=Pct_Typical*100
                     )
                 ) + 
  geom_line(aes(group=School, color=School)) +
  geom_point(color="white", size=8.75) +
  geom_hline(aes(yintercept=80), color="lightgray") +
  geom_text(aes(label=paste(Pct_Typical*100,"%",sep=""), 
                       color=School),
            size=3) +
  scale_color_manual(values = c("#439539", "purple", "#60A2D7", "#C49A6C")) +
  facet_grid(Subject~Grade) +
  theme_bw() + 
  theme(legend.position="bottom") +
  xlab("School Year") +
  ylab("% Meets/Exceeds\nTypical Growth" )  
  
  # return
  p.long
}
```

let's try plotting:

```{r plot, fig.height=6, fig.width=11, dev='svg'}
plot(test_mv_summary)
```

Now that is cooking with gas!!!




