---
title: "MAP Projector"
author: "Chris Haid & Andrew Martin"
date: "October 1, 2014"
output: html_document
runtime: shiny
---
####What is this
__MAP Projector__ uses NWEA's publised norm tables---the [2011 RIT Scale Norms Study Data Files](http://support.nwea.org/node/11789)---to perform a [Markov Process](http://en.wikipedia.org/wiki/Markov_process) simulation of 5,000 potential student trajectories from the RIT scores achieved in the lowest observed grade for the selected subject. (Note: the simulation assumes a fall start date, but not every student's first test event occurs in the fall; the result is that some simulations start too early.  I'll fix that eventually).

You can choose to highlight all simulated paths that end at a given score (say the 75th percentile) in an exiting grade; You can also plot the average path through all the simulated data and through all the highlighted simulations.  You'll notice that a posterior probability of achieving as score at least as the high as the  _highlight cut score_ is given above the visualization.


The projected average paths are a simple linear model that fits $y_{s,t} = \alpha +  \beta\times\log(t) + \epsilon_{s,t}$, where $y_{s,t}$ is a simulated RIT score for sample $s$ at time $t$---an indicator of when a simulated assessment occured and the $\epsilon_{s,t} \sim N(0,1)$ are random errors.
```{r data_prep, echo=FALSE, message=FALSE, warning=FALSE}

# In this chunk we load our packages, data, and do some munging.

# Amazingly, you need only load two packages to makes this bad boy go. 
require(mapvisuals)
require(lubridate)


#load fake MAP Data
data(nweamap)

# Vectorize mapvizier::grade_level_season
gls<-function(x){
  out<-as.vector(mapply(grade_level_season, x, SIMPLIFY=TRUE))
  out
}

# subest for brevity's sake
subjects<-c("Mathematics", "Reading")

map_results<-nweamap %>% 
  filter(MeasurementScale %in% subjects, 
         year(mdy(nweamap$TestStartDate)) >= 2010,
         Grade>=5
         ) %>% mapvizier



students<-map_results$mapData %>% 
  select(StudentID, StudentLastname, StudentFirstname, Grade) %>%
  group_by(StudentID, StudentLastname, StudentFirstname) %>%
  summarize(CurrentGrade=max(Grade)) %>% 
  mutate(StudentName=paste(StudentLastname, StudentFirstname, sep=", ")) %>%
  ungroup %>%
  select(StudentName, StudentID)

stus<-unlist(with(students, split(as.character(StudentID), StudentName)))
```

#### Inputs
```{r inputs, echo=FALSE, messages=FALSE, warnings=FALSE}

# This chunk builds up the input panel.  The call to inputPanel() will creat the silverish
# panel that includes all the input types (in this case, select dropdowns, checkboxes, and a numeric input)
# The corresponding input function (selectInput(), checkboxInput(), numericInput()) save the selected/inputed
# values to an object called input and accessed (via input$inputID) in any of the reacitive functions below 
# used to subset data, generate simulations, and and (re)generate plots 

actionButton("helpButton", label="How This Works")


    renderUI({
      
      if (input$helpButton %% 2 ==0) return()
       div(class="alert",
           "1. Pick a measurement Scale (i.e., a subject to lay people).", br(),
       "2. Pick a student.", br(),
       "3. Select at which grade you'd like simulated data to end.", br(),
       "4. Check any checkbox to draw simulated data and simulated average paths, be patient.", br(),
       "5. Change the final grade value above which simulated scores are highlighted.", br()
       )
      })


h5("Select Student Data")
inputPanel(
  selectInput("ms", label = "Select Measurement Scale",
              choices = subjects, selected = "Reading"),
  
  selectInput("student", label = "Select Student:",
               choices = stus)
  )

h5("Simulated Data Display options")
inputPanel(
  selectInput("end_grade", label = "Select Simulation End Grade",
               choices = c(8:10), selected=10),
  
  checkboxGroupInput("overlays","",
                     choices=c("Show Simulated Data" = "sim_yes", 
                       "Highlight High end score paths" = "hl_yes",
                       "Show average path" = "avg_path_yes",
                       "Show average highlighted path" = "hl_path_yes")),
    
  numericInput(inputId = "ontrack_rit", 
               value=250,
               label = "Highlight simulated scores at or above:",
               min = 195, max=300, step=10)
)


```


```{r reactive_date, echo=FALSE, messages=FALSE, warnings=FALSE}

# This chunk contains all "reactive" elements that aren't directly pasted to a renderX function:  data, 
# variables, and the simulation itself.  Each of these objects is a reactive function: these functions watch 
# the inputs that will change their value say by sugsetting the data by choosing a student or a measurement 
# scale.

#Subset map data by sutdent, and subject, add a grade_level_season measure
plot_data <- reactive({
  plot_data <- map_results$mapData %>% 
    filter(StudentID==as.integer(input$student), 
    MeasurementScale==input$ms) %>%
    mutate(GradeSeason=Grade + gls(Season))
  plot_data 
})

#Get minimum grade from plot_data to start simulation
start_grade <- reactive({
    min(plot_data()$Grade)
})

#Get starting rit form minimum data in map data. used to starting simulation 
start_rit <- reactive({
    start_rit<-dplyr::filter(plot_data(), Grade==min(Grade), 
                             mdy(TestStartDate)==min(mdy(TestStartDate)))
    start_rit$TestRITScore
})

#This runs the simulation with mapvizier::norm_sim().  Notice it receives two reactive values (start_grade()
# and start_rit(), since these are used both here in in the plot) as well as two input values (input$ms and 
# input$end_grade)
sim<-reactive({
  
  sim<-norm_sim(start.grade=start_grade(), 
                start.subject = input$ms,
                end.grade = input$end_grade,
                start.rit = start_rit()
                )
  
   sim<-sim %>% 
    data.frame %>% 
    mutate(Season=ifelse(Season==41,"Fall","Winter"),
           GradeSeason=StartGrade+gls(Season),
           TestRITScore=StartRIT)
  
    select_vars(names(sim), TestRITScore=StartRIT)
  
  sim_high <- filter(sim, GradeSeason==max(GradeSeason)) %>% 
    mutate(High=TestRITScore>=input$ontrack_rit) %>%
    select(ID, High)
  
  sim <- sim %>% left_join(sim_high, by="ID")
  sim
})

proby <- reactive({
  proby <- filter(sim(), GradeSeason==max(GradeSeason)) %>% data.frame %>%
    dplyr::summarize(N=n(), Pr=round(sum(High)/N*100,2))
  proby
})

```

#### Visualization
```{r visualization, echo=FALSE, messages=FALSE, warnings=FALSE}

# This chunk renders the ggplot-based visualization. It adds the simuluated paths and highlighted paths
# via the checkboxInputs above (input$sim_yes and input$hl_yes) as well as the reactive functions sim() and 
# plot_data()


renderText({paste0("Simluated probability that student achieves score higher than ", 
                   input$ontrack_rit, 
                   " in highest grade is ", 
                   proby()$Pr, 
                   "%")})

renderPlot({
  
  # initiats ggplot call using plot_data() reactive function
  withProgress(message="Drawing graph", value=0.1, {
    p<-ggplot(plot_data(), aes(x=GradeSeason, y=TestRITScore))
    incProgress(.2, detail="Initiating plot with student data")
  
  if("sim_yes"%in% input$overlays){ # show simulated futures
    incProgress(.4, detail="Drawing simulated data.")
    
    p<-  p + geom_line(data=sim(), 
              aes(x=GradeSeason, y=TestRITScore, group=ID), 
              color='gray',
              alpha=.025) 
  } 
  if("hl_yes" %in% input$overlays){ # highlight those futues that end at end_grade and exceed ontrack_rit
    incProgress(.1, detail=paste("Highlighting paths exceeding", 
                                  input$ontrack_rit, 
                                  "by grade", input$end_grade))
    
    p <-  p + geom_line(data=filter(sim(), High==TRUE), 
              aes(x=GradeSeason, y=TestRITScore, group=ID), 
              color='lightgreen',
              alpha=.05) 
  } 
  if("avg_path_yes" %in% input$overlays){
    
    incProgress(.1, detail="Calculating and adding average path.")
    
    p <- p + stat_smooth(data=sim(), method=lm, formula=y~log(x), color="#8D8865")
  }
  if("hl_path_yes" %in% input$overlays){
    incProgress(.05, detail="Calculating and adding average highlighted path.")
    p <- p + stat_smooth(data=filter(sim(), High==TRUE), method=lm, formula=y~log(x), color="#439539")
  }
#plot selected students history
  
 
  
p <- p + 
    geom_line(color='blue') +
    theme_bw() + 
    ylab("RIT Score") +
    xlab("Grade") 
  
  
  print(p)

  setProgress(.95, detail="drawing all elements")
  setProgress(1)


    
  })  
    
})

```
```{r datatable, echo=FALSE}
tabsetPanel(
  tabPanel("Simulated Data",
    # This is simply a JavaScript based sortable, searchable, filterable DataTable using the data returned by
    # the sim() reactive function.
    renderDataTable({
      sim()
      })
    ),
    tabPanel("Student Data",
             renderDataTable({
               plot_data()
               })
             )
  )




```




